The impacts of park access on health outcomes: A spatial comprehensive approach

### 2019-03-01 manuscript.Rnw
Add reference and clarify 500 Cities data

\corrauth{Gregory Macfarlane\\
430 Engineering Building\\
Brigham Young University\\
Provo, UT 84602\\
USA}
\begin{abstract}
Recent research has underscored the potential for public green spaces to
influence individual and societal health outcomes, but
empirical measurements of such influences have yielded mixed results to date, with
particular disagreement surrounding how access to parks ought to be defined while
controlling alternate explanations. In this paper we present a comprehensive
measure of park accessibility drawn from random utility choice theory, which avoids
arbitrary assertions of proximity while incorporating potentially numerous
amenities and attributes of both the parks and the population. We apply this
choice-based accessibility measure to Census tract-level health and
socioeconomic data in New York City, paired with geographic parks data and
geolocated position data collected from the Twitter API. Controlling for the
socioeconomic variables and spatially correlated error terms, we show a positive
and significant relationship between park access and physical activity rates, and
a clear negative relationship between park access and obesity rates. Applying
a more traditional analysis considering 1/4 mile acess to parks yields mixed
results.
\end{abstract}

@@

Note that this formulation is an extension of the gravity-based accessibility
statistics, in the same way that the gravity-based trip distribution model
is a specific case of
a destination-choice model \citep{Daly1982}. But the extension is meaningful and
allows us to consider more attributes of the park or the individual beyond size
and impedance. For example, we can aggregate the number of tweets emanating from
a park as a measure of the park's attractiveness beyond its size, and the
utility for the choice maker from Equation \ref{eq:v_sd} then becomes
\begin{equation}\label{eq:v_sdt}
V_{ij} = size_j\lambda_s + distance_{ij}\lambda_d + tweets_j\lambda_t
\end{equation}

@@

The Centers for Disease Control and Prevention makes small-area estimates on
key health indicators available through its 500 Cities data program \citep{CDC5002016}. The models
are estimated on Census sociodemographic data paired with national and local
health surveys, and have been validated at multiple geographic levels
\citep{Wang2018, Wang2017, Zhang2015}. We use two synthetic indicators as our
outcome variables in this study: the share of adults in a Census tract who are
obese, and the share of adults who participate in no leisure-time physical activity.
To improve clarity in our interpretation, we use the complement of the second variable ---
the share of tract adults who participate in {\em some} physical activity ---
even if the amount is not sufficient to affect overall health. Both indicators
are estimated for the year 2016.

>>>>

\corrauth{Gregory Macfarlane\\
430 Engineering Building\\
Brigham Young University\\
Provo, UT 84602\\
USA}
\begin{abstract}
Recent research has underscored the potential for public green spaces to
influence individual and societal health outcomes, but
empirical measurements of such influences have yielded mixed results to date, with
particular disagreement surrounding how access to parks ought to be defined while
controlling alternate explanations. In this paper we present a comprehensive
measure of park accessibility drawn from random utility choice theory, which avoids
arbitrary assertions of proximity while incorporating potentially numerous
amenities and attributes of both the parks and the population. We apply this
choice-based accessibility measure to correlate Census tract-level obesity and
physical activity rate estimates from the Centers for Disease Control and
Prevention's 500 Cities project with tract-level American Community Survey
socioeconomic data in New York City, paired with geographic parks data from New
York City and geolocated tweet data collected from the Twitter API as a metric of
a park's activity levels. Controlling for the socioeconomic variables and
spatially correlated error terms, we show a positive and significant
relationship between park access and physical activity rates, and a clear and
significant negative relationship between park access and obesity rates.
Applying a more traditional analysis considering 1/4 mile acess to parks yields
mixed results and cannot accommodate features such as park size and twitter
activity.
\end{abstract}

@@

Note that this formulation is an extension of gravity-based accessibility
statistics, in the same way that the gravity-based trip distribution model
is a specific case of
a destination-choice model \citep{Daly1982}. But the extension is meaningful and
allows us to consider more attributes of the park or the individual beyond size
and impedance. For example, we can aggregate the number of tweets emanating from
a park as a measure of the park's attractiveness beyond its size, and the
utility for the choice maker from Equation \ref{eq:v_sd} then becomes
\begin{equation}\label{eq:v_sdt}
V_{ij} = size_j\lambda_s + distance_{ij}\lambda_d + tweets_j\lambda_t
\end{equation}

@@

The Centers for Disease Control and Prevention makes small-area estimates on
key health indicators available through its 500 Cities data program \citep{CDC5002016}.
The indicators are multilevel aggregations and imputations of BRFSS responses
\citep{Wang2018, Wang2017}, and have been recently used to study the tract-level
link between gentrification and urban health \citep{Gibbons2018}. We use two
indicators as our dependent variables in this study: the share of adults
in a Census tract who are obese, and the share of adults who participate in no
leisure-time physical
activity. To improve clarity in our interpretation, we use the complement of the
second variable --- the share of tract adults who participate in {\em some}
physical activity --- even if the amount is not sufficient to affect overall
health. Both indicators are estimated for the year 2016.
